---
title: "EDA Report for Apex Trust Dataset"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 5)

# Load necessary libraries
library(tidyverse)
library(gridExtra)
library(Amelia)
library(caret)
library(psych)
library(vip)

# Import dataset
Apex_Trust_Dataset <- read.csv("C:/Users/amdari/projects/Credit Risk Assesment/Apex_Trust_Dataset.csv")

# Duplicate dataset for EDA
eda_data <- Apex_Trust_Dataset

# Assign categorical columns to an object and convert to factors
cat_cols <- c('status', 'duration', 'credit_history', 'purpose', 'amount', 'savings',
              'employment_duration', 'installment_rate', 'other_debtors', 'present_residence',
              'property', 'age', 'other_installment_plans', 'housing', 'number_credits',
              'job', 'people_liable', 'telephone', 'foreign_worker', 'credit_risk')
eda_data[, cat_cols] <- lapply(eda_data[, cat_cols], factor)

# Data Transformation
# Rename variable values for better readability
eda_data <- eda_data |> mutate(credit_risk = ifelse(credit_risk == 0, 'bad', 'good'))

# Transform 'status' and 'savings' columns
eda_data$status <- factor(ifelse(eda_data$status == 1, 'no checking account', 
                                  ifelse(eda_data$status == 2, '<0 USD', 
                                         ifelse(eda_data$status == 3, '0 USD >= & < 200 usd', '>200 USD'))),
                          levels = c('no checking account','<0 USD','0 USD >= & < 200 usd','>200 USD'))
eda_data$savings <- factor(eda_data$savings, levels = c('unknown/ no savings account', '<100 USD', 
                                                        '100 >= & < 500 usd', '500 >= & 1000 USD', '>=1000 USD'))

# Data Exploration
# View the structure and summary statistics of the transformed data.
```{r test-chunk, echo=TRUE}
summary(eda_data)

# Exploratory Data Analysis (EDA)
# Distribution of Response Variable
credit_risk_dist <- ggplot(eda_data, aes(x = credit_risk)) +
  geom_bar(width = 0.25, fill = 'darkgreen') +
  theme_minimal() +
  labs(x = 'Credit Risk', y = 'Count', title = "Distribution of Response Variable") +
  theme(plot.title = element_text(size = 17, family = "sans", hjust = 0.5))
print(credit_risk_dist)

# Create a separate data frame for EDA
df2 <- eda_data |> mutate(credit_risk = as.factor(ifelse(credit_risk == 'bad', 0, 1)))

# Define a function to create bar plots for categorical variables
create_bar_plot <- function(data, var_name) {
  ggplot(data, aes_string(x = var_name)) +
    geom_bar(aes(fill = after_stat(count)), color = "black") +
    facet_wrap(~credit_risk, nrow = 2) +
    scale_fill_gradient("Count", low = "lightgreen", high = "darkgreen") +
    theme_minimal() +
    labs(title = paste("Distribution of", var_name))
}

# Create plots for each categorical variable
print(create_bar_plot(df2, "status"))
print(create_bar_plot(df2, "duration"))
print(create_bar_plot(df2, "age"))
print(create_bar_plot(df2, "credit_history"))
print(create_bar_plot(df2, "job"))
print(create_bar_plot(df2, "number_credits"))
print(create_bar_plot(df2, "other_installment_plans"))
print(create_bar_plot(df2, "savings"))
print(create_bar_plot(df2, "amount"))

# Data Preprocessing
# Check for missing values and visualize with missmap.
missmap(Apex_Trust_Dataset, main = "Missing Values Map")

# Model Development
# Split the data into training and test sets
set.seed(1234)
ind <- sample(2, nrow(Apex_Trust_Dataset), replace = TRUE, prob = c(0.7, 0.3))
train <- Apex_Trust_Dataset[ind == 1,]
test <- Apex_Trust_Dataset[ind == 2,]

# Ensure 'credit_risk' is a factor in training and test sets
train$credit_risk <- factor(train$credit_risk, levels = c(0, 1))
test$credit_risk <- factor(test$credit_risk, levels = c(0, 1))

# Build a logistic regression model
m1 <- glm(credit_risk ~ ., data = train, family = 'binomial')
print(summary(m1))

# Model Evaluation
# Predictions on training data
p1 <- predict(m1, train, type = 'response')
pred1 <- ifelse(p1 > 0.5, 1, 0)
cmlg1 <- confusionMatrix(factor(pred1), factor(train$credit_risk), positive = '1')
print(cmlg1)

# Predictions on test data
p2 <- predict(m1, test, type = 'response')
pred2 <- ifelse(p2 > 0.5, 1, 0)
cmlg2 <- confusionMatrix(factor(pred2), factor(test$credit_risk), positive = '1')
print(cmlg2)

# Feature Importance Visualization
# Display the most important features in the model
m1 %>% vip(num_features = 20, geom = "point", aesthetics = list(size = 2, color = "#3C6255")) +
  theme_minimal(base_size = 18) +
  labs(title = "Logistic Regression: Feature Importance")
